Determinism, Complexity, and Predictability in Computer Performance

### 2014-08-04 concl.tex
maybe done

\alert{add FW stuff from section VI}

@@

In this paper, we have argued that redundancy is a useful definition
of the inherent predictability of an empirical time series.  To
operationalize that definition, we use an approximation of the
Kolmogorov-Sinai entropy, estimated using a weighted version of the
permutation entropy of~\cite{bandt2002per}.  This WPE technique---an
ordinal calculation of forward information transfer in a time
series---is ideal for our purposes because it works with real-valued
data and is known to converge to the true entropy value. Using a
variety of forecast models and more than 150 time-series data sets
from experiments and simulations, we have shown that prediction
accuracy is indeed correlated with weighted permutation entropy: the
higher the WPE, in general, the higher the prediction error.  The
relationship is roughly logarithmic, which makes theoretical sense,
given the nature of WPE, predictability, and MASE.

@@

The curves and shaded regions in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} generalize and operationalize the
discussion in the previous paragraph.  These geometric features are a
preliminary, but potentially useful, empirical heurists for knowing
when a model is not well-matched to the task at hand: a point that is
below and/or to the right of the shaded regions on a plot like
Figure~\ref{fig:wpe_vs_mase_best} indicates the time series has more
predictive structure than the forecast model can capture and
exploit---and that one would be well advised to try another method.

These curves were determined empirically using a specific error metric
and a finite set of forecast methods and time series traces.  If one
uses a different error metric, the geometry---and the heuristic---will
no longer apply.  And while the methods and traces were chosen to be
representative of the practice, they are of course not completely
comprehensive.  It is certainly possible, for instance, that the
nonlinear dynamics of computer performance is subtly different from
the nonlinear dynamics of other systems.  Our preliminary results on
other systems (H\'enon, Lorenz, a white-noise process, SFI ``A'',
transformations of the computer performance data) lead us to believe
that our results will generalize beyond the examples described in this
paper.  We are in the process of following up on that exploration with
a broader study of data, forecast methods, and error metrics.

@@

Nonstationarity is a serious challenge in any time-series modeling
problem.  Any regime shift that causes a change in the predictive
structure between the training signal and the test signal, for
instance, will skew the MASE score (and the points on the curve).
This can destroy the utility of our heuristic.  \alert{Conversely,
  though,} a MASE score of a random-walk prediction that is larger
than 1---as mentioned at end of the previous section---can potentially
be a good indicator that a signal is nonstationary.  Detecting regime
shifts---and adapting prediction models accordingly---is an important
area of future work.  Indeed, one of the first applications of
permutation entropy was to recognize the regime shift in brainwave
data that occurs when someone has a seizure~\cite{cao2004det}.  Recall
that the signal in Figure~\ref{fig:svd-ts-colored} was especially
useful for the study in this paper because it contained a number of
different regimes.  We segmented this signal visually, but one could
imagine using permutation entropy to do so instead.  Automating
regime-shift detection would be an important step towards a fully
adaptive modeling strategy, where old models are discarded and new
ones are rebuilt whenever the time series enters a new regime.  WPE
would be particularly powerful in this scenario, as its value can not
only help with regime-shift detection, but also suggest what kind of
model might work well in each new regime.  Of particular interest
would be the class of so-called \emph{hybrid systems}~\cite{hybrid},
which exhibit discrete transitions between different continuous
regimes---e.g., a lathe that has an intermittent instability or
traffic at an internet router, whose characteristic normal traffic
patterns shift radically during an attack.  Effective modeling and
prediction of these kinds of systems is quite difficult; doing so
adaptively and automatically---in the manner that is alluded to at the
end of the previous paragraph---would be an interesting challenge.

>>>>

@@

In this paper, we have argued that redundancy is a useful proxy for
the inherent predictability of an empirical time series.  To
operationalize that relationship, we use an approximation of the
Kolmogorov-Sinai entropy, estimated using a weighted version of the
permutation entropy of~\cite{bandt2002per}.  This WPE technique---an
ordinal calculation of forward information transfer in a time
series---is ideal for our purposes because it works with real-valued
data and is known to converge to the true entropy value. Using a
variety of forecast models and more than 150 time-series data sets
from experiments and simulations, we have shown that prediction
accuracy is indeed correlated with weighted permutation entropy: the
higher the WPE, in general, the higher the prediction error.  The
relationship is roughly logarithmic, which makes theoretical sense,
given the nature of WPE, predictability, and MASE.

@@

The curves and shaded regions in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} generalize and operationalize the
discussion in the previous paragraph.  These geometric features are a
preliminary, but potentially useful, heuristic for knowing when a
model is not well-matched to the task at hand: a point that is below
and/or to the right of the shaded regions on a plot like
Figure~\ref{fig:wpe_vs_mase_all} indicates that the time series has
more predictive structure than the forecast model can capture and
exploit---and that one would be well advised to try another method.

These curves were determined empirically using a specific error metric
and a finite set of forecast methods and time-series traces.  If one
uses a different error metric, the geometry of the heuristic will be
different---and may not even make sense, if one uses a metric that
does not support comparison across different time series.  And while
the methods and traces used in this study were chosen to be
representative of the practice, they are of course not completely
comprehensive.  It is certainly possible, for instance, that the
nonlinear dynamics of computer performance is subtly different from
the nonlinear dynamics of other systems.  Our preliminary results on
other systems (H\'enon, Lorenz, a white-noise process, SFI ``A'',
transformations of the computer performance data) lead us to believe
that our results will generalize beyond the examples described in this
paper.  We are in the process of following up on that exploration with
a broader study of data, forecast methods, and error metrics.

@@

Nonstationarity is a serious challenge in any time-series modeling
problem.  Any regime shift that causes a change in the predictive
structure between the training signal and the test signal, for
instance, will skew the MASE score (and the points on the curve),
obviating the utility of our heuristic.  \alert{Conversely, though,} a
MASE score of a random-walk prediction that is larger than 1---as
mentioned at end of the previous section---could potentially be a good
indicator that a signal is nonstationary.  Detecting regime
shifts---and adapting prediction models accordingly---is an important
area of future work.  Indeed, one of the first applications of
permutation entropy was to recognize the regime shift in brainwave
data that occurs when someone has a seizure~\cite{cao2004det}.  Recall
that the signal in Figure~\ref{fig:svd-ts-colored} was especially
useful for the study in this paper because it contained a number of
different regimes.  We segmented this signal visually, but one could
imagine using permutation entropy to do so instead.  Automating
regime-shift detection would be an important step towards a fully
adaptive modeling strategy, where old models are discarded and new
ones are rebuilt whenever the time series enters a new regime.  WPE
would be particularly powerful in this scenario, as its value can not
only help with regime-shift detection, but also suggest what kind of
model might work well in each new regime.  Of particular interest
would be the class of so-called \emph{hybrid systems}~\cite{hybrid},
which exhibit discrete transitions between different continuous
regimes---e.g., a lathe that has an intermittent instability or
traffic at an internet router, whose characteristic normal traffic
patterns shift radically during an attack.  Effective modeling and
prediction of these kinds of systems is quite difficult; doing so
adaptively and automatically---in the manner that is alluded to at the
end of the previous paragraph---would be an interesting challenge.
